---
title: "miRNA_targets"
author: "Francis Leblanc"
date: "20/11/2020"
output:   
  html_document:
    toc: yes
    toc_float: yes
    number_sections: true
    code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```



```{r}
library(rio)
library(dplyr)
library(miRNAtap)
library(org.Hs.eg.db)
```


# Investigating overlaping targets predicted by 5 databes with DE genes 
\n


**MiRNAtap** aggregates targets from 5 most commonly cited prediction algorithms:

* DIANA (Maragkakis et al., 2011),
* Miranda (Enright et al., 2003), 
* PicTar (Lallet al., 2006), 
* TargetScan (Friedman et al., 2009), and 
* miRDB (Wong and Wang,2015).

#### The results are given in searchable tables
\n
The *SYMBOL* column is human gene names. 
If no gene names was found for that human gene in CanFam3.1, no correlations can be obtained and correlation column is empty (this is the case for ~50 of the targets called with these miRNAs)
Navigate the table to page 6 to start seing the correlations or use sorting on columns.
The Protein tables are at the end of the document.
\n

# Overlap of targets and DE mRNAs

## dlk1_dio3

```{r}
all_res <- import("../mRNA/data/final/all.FB.corrected_expressions.RDS")
miRNAs <- all_res %>% filter(Dataset == "miRNA" & chromosome_name == 8 & CTLvsAVBpadj < 0.01)

# subset miRNA DE from dlk1-dio3 locus
dlk1_dio3 <- miRNAs %>% pull(Gene) 
dlk1_dio3 <- dlk1_dio3[-grep("ENS",dlk1_dio3)]
dlk1_dio3 <- gsub("MIR", "mir-", dlk1_dio3)

df <- data.frame(source_1=integer(),
                 source_2=integer(),
                 source_3=integer(),
                 source_4=integer(),
                 source_5=integer(),
                 rank_product=integer(),
                 rank_final=integer(),
                 miRNA = character(),
                 ENTREZ = character())

# Query miRNA target databases
for (i in dlk1_dio3) {
  df1 <- as.data.frame(getPredictedTargets(i, species = 'hsa', method = 'geom', min_src = 3))
  df1$miRNA <- i
  df1$ENTREZ <- row.names(df1)
  df <- dplyr::bind_rows(df,df1)}

# Get miRNA
Symb_Entr <- select(org.Hs.eg.db, df$ENTREZ, "SYMBOL", "ENTREZID")
df$SYMBOL <- Symb_Entr$SYMBOL

# Keep DE mRNAs only
DE_mRNA_CTLvsAVB <- all_res %>% filter(Dataset == "mRNA" & CTLvsAVBpadj < 0.01) %>% pull(Gene)
DE_targets_CTLvsAVB <- df[df$SYMBOL %in% DE_mRNA_CTLvsAVB,]

# Rename miRNAs to match original data
DE_targets_CTLvsAVB$miRNA_Gene <- gsub("mir-", "MIR", DE_targets_CTLvsAVB$miRNA)
```


# Filter based on negative correlation

```{r}
DE_targets.all <- DE_targets_CTLvsAVB
DE_targets.all$cors <- NA

#extract pearson correlations
for(i in unique(DE_targets_CTLvsAVB$miRNA_Gene)) {
  #get targets 
  targets <- DE_targets_CTLvsAVB$SYMBOL[DE_targets_CTLvsAVB$miRNA_Gene == i]
  targets <- targets[targets != "" & !(is.na(targets))]
  
  #get expression values from mRNA for the targets of that miRNA
  mRNA.x <- all_res[match(targets, all_res$Gene),c(16,18:35)]
  colnames(mRNA.x)[1] <- "Gene"

  #add expression values from the miRNA and transpose
  mRNA.x <- rbind(miRNAs[miRNAs$Gene == i,c(1,18:35)], mRNA.x)
  mRNA.x.m <- data.matrix(mRNA.x[,2:19], rownames.force = NA)
  mRNA.x.m <- t(mRNA.x.m)
  
  #get correlation values 
  cors <- data.frame(cor(mRNA.x.m))
  colnames(cors) <- mRNA.x$Gene
  
  #make a dataframe to merge previous informations
  cors_f <- data.frame(miRNA_Gene = colnames(cors)[1],
                     SYMBOL = as.character(colnames(cors)[2:length(colnames(cors))]),
                     cor.tmp = cors[2:length(colnames(cors)),1])

  DE_targets.all <- merge(DE_targets.all, cors_f, all.x = T) 
  
  #add i's targets correlations to cors
  DE_targets.all$cors[is.na(DE_targets.all$cors)] <- DE_targets.all$cor.tmp[is.na(DE_targets.all$cors)]
  
  #remove tmp col
  DE_targets.all <- subset(DE_targets.all, select = -cor.tmp)

}

# filter based on negative correlation
DE_targets.filt <- DE_targets.all %>% filter(cors < -0.5)


export(DE_targets.filt, "data/final/dlk1_dio3_targets.RDS")
```

